---
title: "Assignment 6"
Author: "Fizza Fatima"
output: html_document
date: "2025-10-07"
---

```{r setup, include=FALSE}
# Load necessary libraries
library(ggplot2)  # For data visualization
library(dplyr)    # For data manipulation
library(lubridate) # For date handling
```


```{r}
# Load the dataset
data_path <- "C:/Assignment 6/Supermart Grocery Sales - Retail Analytics Dataset.csv"
sales_data <- read.csv(data_path)

# View the first few rows of the data
head(sales_data)
```
It is the step that loads the dataset in the specified file path (.csv) to the sales-data variable as a data frame.
```{r}
# Data Cleaning
# Convert 'Order Date' to Date format
sales_data$Order.Date <- dmy(sales_data$Order.Date)  # 'dmy' is used for dates like "DD-MM-YYYY"

# Remove rows with missing data
sales_data <- sales_data %>% na.omit()

# Create a new column for Profit-to-Sales ratio
sales_data$Profit_To_Sales <- sales_data$Profit / sales_data$Sales
```
dmy() converts the Order.Date column from text to Date format (DD-MM-YYYY).
na.omit() removes rows with any missing values.
```{r}
# Total Sales by Category
# Check column names
colnames(sales_data)

# Check the structure of the dataset
str(sales_data)

# Convert 'Sales' to numeric (if it's not already)
sales_data$Sales <- as.numeric(sales_data$Sales)

# Convert 'Category' to factor (if it's not already)
sales_data$Category <- as.factor(sales_data$Category)

# Check for missing values
sum(is.na(sales_data$Sales))  # Number of missing values in Sales
sum(is.na(sales_data$Category))  # Number of missing values in Category

# Remove rows with missing Sales or Category values
sales_data <- sales_data %>% filter(!is.na(Sales) & !is.na(Category))

# Aggregate sales by category
category_sales <- sales_data %>%
  group_by(Category) %>%
  summarise(Total_Sales = sum(Sales, na.rm = TRUE))

# Check the aggregated data
head(category_sales)


# Bar chart: Total Sales by Category
ggplot(data = category_sales, aes(x = Category, y = Total_Sales, fill = Category)) +
  geom_bar(stat = "identity") + 
  labs(title = "Total Sales by Category", x = "Category", y = "Sales") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate category names

```
ggplot() sets up the plot, specifying Category as the x-axis and Total_Sales as the y-axis.
geom_bar(stat = "identity") creates a bar chart where the height of each bar is the Total_Sales for each category.
labs() adds the title and axis labels.
theme(axis.text.x = element_text(angle = 45, hjust = 1)) rotates the x-axis labels for better readability.
```{r}
# Scatter plot: Sales vs Profit
ggplot(data = sales_data, aes(x = Sales, y = Profit, color = Category)) +
  geom_point() +
  labs(title = "Sales vs Profit", x = "Sales", y = "Profit") +
  theme_minimal()

```
ggplot() sets up the plot, specifying Sales as the x-axis and Profit as the y-axis.
geom_point() creates a scatter plot, with each point representing a sales-profit pair.
color = Category assigns different colors to each category for better differentiation.
labs() adds the title and axis labels.
theme_minimal() gives the plot a minimalistic style.
